---
title: "Data Taming Assignment 1"
author: "Dongju Ma"
date: "12/06/2024"
output:
  pdf_document: default
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

```{r }
#Load the required packages
library(tidyverse)
library(inspectdf)
```

# Q1. Loading the data

```{r}
# Your student number goes here
ysn = 1942340
# Calculate your student number modulo 3
filenum <- ysn %% 3
filenum
filename <- paste0("./data/afl_",filenum,".csv")
filename

# Read in the data
afl<-read_csv("./data/afl_2.csv")
# Display the first 10 lines of the data
head(afl,10)
```

# Q2. The dimensions of the data set

```{r}
#Use dim to show the numbers of rows and columns
dim(afl)
```

The data set has 18 rows and 24 columns.


# Q3. Random permutation of the rows

```{r}
# Set the random seed
set.seed(1942340)
# Use sample_n to get the random permutation of the rows
afl1<-sample_n(afl,18,replace = FALSE)
afl1
```

# Q4. Adding an extra column of row numbers
```{r}
# Use mutate to add a column at the far right of the data set
afl1<-mutate(afl1,RowNum=c(1:18))
# Then use relocate to move the new column to the far left
afl1<-relocate(afl1,"RowNum", .before = Team)
afl1
```

# Q5 Data cleaning
## Q5(a)
```{r}
# Use filter to extract the rows without text data.
afl1<-filter(afl1,Team!="testX1")
# Make sure the row numbers are updated
afl1<-mutate(afl1,RowNum=c(1:17))
afl1
```

## Q5(b)
```{r}
# Change Team name "Adelaide" to "Port Adelaide"
afl1[9,]$Team<-str_replace(afl1[9,]$Team,"Adelaide","Port Adelaide")
# Change Team name "Melbourne" to "North Melbourne"
afl1[14,]$Team<-str_replace(afl1[14,]$Team,"Melbourne","North Melbourne")
# Change State "Queensld" to "QLD"
afl1[4,]$State<-str_replace(afl1[4,]$State,"Queensld","QLD")
# Change State "New South Wales" to "SA"
afl1[9,]$State<-str_replace(afl1[9,]$State,"New South Wales","SA")
# Change State "bictoria" to "VIC"
afl1[15,]$State<-str_replace(afl1[15,]$State,"bictoria","VIC")
afl1
```

## Q5(c)
```{r}
# Use arrange to sort the tibble by team name
afl1<-arrange(afl1,Team)
afl1
```

# Q6
## Q6(a)
```{r}
# Use gather to convert the data set to long form
afl1<- gather(afl1,key = "round",value = "details",'Round01':'Round22')
afl1
```
## Q6(b)
```{r}
# Use sting replace to remove all the "Round" string in column round
afl1$round<-str_replace(afl1$round,"Round","")
afl1
```
## Q6(c)
```{r}
# Judge is away in details column, and rename the result column 1 into home
afl1<-afl1 %>%
  mutate("home"=is.na(str_match(afl1$details,"away"))[,1])
afl1
```
## Q6(d)
```{r}
# Dig the numbers by str_match and put the result into column goals and column behinds
afl1<-mutate(afl1,goals=str_match(afl1$details,"(\\d+) goals and (\\d+)")[,2])
afl1<-mutate(afl1,behinds=str_match(afl1$details,"(\\d+) goals and (\\d+)")[,3])
afl1
```
## Q6(e)
```{r}
# Delete the details column
afl1<-mutate(afl1,details=NULL)
afl1
```
## Q6(f)
```{r}
# Add the TidyRowNum column right next to the origin RowNum
afl1<-mutate(afl1,TidyRowNum=(1:374), .after=RowNum)
afl1
```




# Q7. Identifying data types

* Row Num: Categorical Ordinal. The numbers represent the teams and round status is home or away. For example number 1 indicates team Carlton's away games. 

* Tidy Row Num: Categorical Ordinal. The tidy row numbers are integers indicate the order of this data set.

* Team: Categorical Nominal. They are the names of teams in AFL.

* State: Categorical Nominal.. They are the names of the states.

* Round: Categorical Nominal. The characters represents the rounds in the match season, which is in the range of 01 to 22.

* home: Categorical Nominal. There are only two categories in this variables, TRUE means the game is home and FALSE means away.

* goals: Quantitative Discrete. The numbers are integers represent the goals' points in each game and they can be really huge theoretically.

* behinds: Quantitative Discrete. The numbers are integers represent the points in behinds and they can be really huge theoretically.

# Q8. Taming the data
```{r}
# Change the blank spaces in Team into "_"
afl1$Team<-str_replace(afl1$Team," ","_")
afl1
# Change the number characters into integers
afl1$round<-as.integer(afl1$round)
afl1$goals<-as.integer(afl1$goals)
afl1$behinds<-as.integer(afl1$behinds)
# Check if there is any NA
inspect_na(afl1)
```

# Q9 Set the new data set
```{r}
set.seed(1942340)
afl2<-sample_n(afl1,200)
afl2
```


# Q10(a) Insert two new columns
```{r}
# Calculate the score and accuracy and insert the new colunms
afl2<-mutate(afl2,score=goals*6+behinds)
afl2<-mutate(afl2,accuracy=goals/(goals+behinds))
afl2
```
The score variable is Quantitative Discrete while the accuracy variable is Quantitative Continuous.
The score's type is incorrect, it should be integers and the accuracy's is correct.
```{r}
afl2$score<-as.integer(afl2$score)
afl2
```
# Q10(b) 
```{r}
summarise(group_by(afl2,Team),mean_score=mean(score))
summarise(group_by(afl2,Team),mean_accuracy=mean(accuracy))
```
* i. Fremantle	104.50000	
* ii. Richmond	75.33333
* iii. Fremantle	0.5674431	
* iv. Collingwood	0.4771722	

# Q11(a)
```{r}
ggplot(afl2,aes(home,score,fill=home))+
  geom_boxplot()
```

# Q11(b)
```{r}
ggplot(afl2,aes(home,accuracy,fill=home))+
  geom_boxplot()
```
The home games have better probabilities to win more scores but the accuracy between home and away is very close. With the graph we can see the average line of home score is higher and the top and bottom is higher than away's. But when it comes to the accuracy graph their posistion is much closer.
# Q12
```{r}
afl_home<-filter(afl2,home==TRUE)
afl_away<-filter(afl2,home==FALSE)
afl_home
afl_away
```
# Q13
```{r}
inspect_num(afl_home)
inspect_num(afl_away)
```
The average score of home games is 94.0510204 while the average accuracy is 0.5218593. Also the average score of away games 86.9803922	while the average accuracy is 86.9803922. 

The data does support the claim.

# Q14
```{r}
ggplot(afl_home, aes(x = accuracy, y = score)) + 
  geom_point() + 
  geom_smooth(method="lm")
ggplot(afl_away, aes(x = accuracy, y = score)) + 
  geom_point() + 
  geom_smooth(method="lm")
```
The calculation of score is to multiple goals numbers with 6 and behinds just 1 time and the accuracy represent the proportion of goals, which infers that with higher accuracy come to higher goals. And the higher goals change into higher scores.
So I choose the accuracy to be the independent variable and the score to be predictor.

# Q15
As the scatter plots shown, the higher accuracy come to higher scores. And it is similar for both home and away teams.
